---
title: "Chapter 2 - Data Visualization with ggplot2"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

```{r}
mpg
```

Create a plot of a car's engine siza (displ) and its fuel efficiency (hwy). I'm going to be using the pipe operator %ª% because it improves readability even tough it is introduced in later chapters.
```{r}
mpg %>%
  ggplot() + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

ggplot create a coordinate system. You can add layers to get exactly the graphic you want. Ggplot when run without any additional code creates a blank rectangle.
```{r}
ggplot()
```
  *geom_point* creates a scatterplot
  *mapping* is used to tell ggplot how to graph/visualize the variables you are interested in. It always goes in had with *aes*. For a scatterplot you must indicate within *aes* the x and y variables within the data.
  
  
  ## Graphing Template

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
  
Using the pipe operator

<DATA> %>% 
  ggplot() +
    <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
    
    
Exercises

 Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```
 

How many rows are in mpg? How many columns?
```{r}
dim(mpg)
```


What does the drv variable describe? Read the help for ?mpg to find out.
f = front-wheel drive, r = rear wheel drive, 4 = 4wd


Make a scatterplot of hwy vs cyl.
```{r}
mpg %>%
  ggplot() +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
It's not usefull because they are both categorical variables.
```{r}
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = class, y = drv))
```
 
 Asthetic mappings
 An aesthetic is a visual property of the objects in your plot. 
 Asthetics include the shape, color or sizes of the shapes in the plot. When talking about asthetics, we talk about levels instead of values.
 
 You can assign diferent levels to a variable in the data.
 
 Here you can see the mapping of colors. You can map the colors of your points to variables in the data
```{r}
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
 
 Here you can see the mapping of sizes to the class variable
```{r}
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
 
 alpha controls the transparency
```{r}
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```
 
 shape of the points
```{r}
mpg %>% 
  ggplot() + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
 
 x and y are asthetics. Visual properties that you can map to variables to display information about the data
 In my own words, just like class can make color appear depending on the variable assigned, so can x and y make a plot appear by graphing it in the cartesian plane.
 
 Ggplot automates a lot of things after you map it to a variable. For x-y points, it creates an axis line, with color, size, shape it creates a legend.
 
 
 You can also set asthetic properties manually
```{r}
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy, color = "cyan"))
```
 
 